\pkg{evtree}: Evolutionary Learning of Globally Optimal Classification and Regression Trees in \proglang{R}

### 2011-09-08 pkg/evtree/inst/doc/evtree.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@208 255f4249-2263-4417-b1ea-8498779e27a7
Avoid using T/F, use TRUE/FALSE instead.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@222 255f4249-2263-4417-b1ea-8498779e27a7

\begin{table}
\begin{center}
\begin{tabular}{lrrrrrr}
\toprule
\textsc{Data-set} & \textsc{Instances} &  \multicolumn{5}{l}{\textsc{Attributes}}   \\ 
                  &                    &  binary              & nominal          &  ordered      &  metric     & classes \\ 
\midrule
Glass Identification$^\#$  &   214   &  - &  - &  - & 9 &  6      	 \\
Statlog Heart*         &   270   &  3 &   3 & 1 & 6 &   2     	 \\
Ionosphere$^\#$ &   351   &  2 &  - &  - & 32 & 2  	             \\
Musk$^+$  &  476  & - &  - &  - & 166 & 2                     	 \\
Breast Cancer Database$^\#$ &   685   &  - &  4 & 5 &  - &  2  	 \\
Pima Indians Diabetes$^\#$ &   768   &  - &   - & - & 8 &  2      	 \\
Vowel$^\#$  &  990  & - &  1 &  - & 9 & 11     	                     \\
Statlog German Credit* &  1000   &  2 &  10 & 1 & 7 &  2      	 \\
Contraceptive Method* &   1437   & 3 &  - &  4 &  2 &  3  	     \\
DNA$^\#$  &  3186  & 180 &  - &  - & - & 3   	                       \\
Spam$^+$  &  4601  & - &  - &  - & 57 & 2                      	 \\
MAGIC Gamma Telescope*  &  19020  & - &  - &  - & 10 & 2     	   \\
Servo$^\#$ & 167 &  - & 4 & - &  -  &  -											   \\
Boston Housing$^\#$ & 506 & 1 & - & - &  12  & -                 \\
MEL0101$^\diamond$ &   875 &  1 & 4 & 1 &  108  & -              \\
HDG0202$^\diamond$ &  3933 &  1 & 7 & 1 &   46  & -              \\
HDG0502$^\diamond$ & 8251 &  1 & 7 & 1 &   91  & -               \\
\bottomrule
\end{tabular}
\end{center}
\caption{Description of the evaluated benchmark data-sets. The data-sets marked with a $(^*)$ are from the UCI machine learning repository \citep{Frank+Asuncion:2010}. Data-sets marked with a $(^\#)$ and $(^{+})$ are from the \proglang{R} packages \pkg{mlbench} \citep{mlbench} and \pkg{kernlab} \citep{kernlab}. The $3$ real-world data-sets from the Austrian DRG system are marked with a $(^\diamond)$. } 
\label{tab:dataDescription}
\end{table}

@@

preprocess <- function(d, dname = "datasetname", isclassification = TRUE){
    if(isclassification == TRUE)
        d[,1:3] <- 1-d[,1:3]
    d <- as.data.frame(d)
    colnames(d) <- c("evtree", "rpart", "ctree","evtree", "rpart", "ctree")
    d[,2] <- d[,2] / d[,1] *100
    d[,3] <- d[,3] / d[,1] *100
    d[,1] <- d[,1] / d[,1] *100
    d[,5] <- d[,5] / d[,4] *100
    d[,6] <- d[,6] / d[,4] *100
    d[,4] <- d[,4] / d[,4] *100
    x <- d[ ,1:3]
    y <- d[ ,4:6]
    x <- reshape(x, idvar="samp", times=names(x), timevar = "alg",varying= list(names(x)), direction="long")
    y <- reshape(y, idvar="samp", times=names(y), timevar = "alg",varying= list(names(y)), direction="long")
    x$ds <- dname
    y$ds <- dname
    return( cbind(x,y) )
}

r <- preprocess(d = rglass, dname = "Glass Identification", isclassification = T)
r <- rbind(r, preprocess(d = rheart, dname = "Statlog Heart", isclassification = T))
r <- rbind(r, preprocess(d = rionosphere, dname = "Ionosphere", isclassification = T))
r <- rbind(r, preprocess(d = rmusk, dname = "Musk", isclassification = T))
r <- rbind(r, preprocess(d = rbreastcancer, dname = "Breast Cancer Database", isclassification = T))
r <- rbind(r, preprocess(d = rpima, dname = "Pima Indians Diabetes", isclassification = T))
r <- rbind(r, preprocess(d = rvowel, dname = "Vowel", isclassification = T))
r <- rbind(r, preprocess(d = rcredit, dname = "Statlog German Credit", isclassification = T))
r <- rbind(r, preprocess(d = rcontraceptive, dname = "Contraceptive Method", isclassification = T))
r <- rbind(r, preprocess(d = rdna, dname = "DNA", isclassification = T))
r <- rbind(r, preprocess(d = rspam, dname = "Spam", isclassification = T))
r <- rbind(r, preprocess(d = rmagicgamma, dname = "Magic Gamma Telescope", isclassification = T))
r <- rbind(r, preprocess(d = rservo, dname = "Servo", isclassification = F))
r <- rbind(r, preprocess(d = rbostonhousing, dname = "Boston Housing", isclassification = F))
r <- rbind(r, preprocess(d = rmel0101, dname = "MEL0101", isclassification = F))
r <- rbind(r, preprocess(d = rhdg0202, dname = "HDG0202", isclassification = F))
r <- rbind(r, preprocess(d = rhdg0502, dname = "HDG0502", isclassification = F))

@@

plotd <- function(alg){
    par(las=1, cex=1.5, cex.lab=1.3, mfrow= c(1,2), omd=c(0.16,1,0,1), mar=c(5, 0, 1, 2) + 0.1)
    dummy <- c(1:17)
    names(dummy) <- uniquedatasetlabels
    stats <- cstats(ra, alg)
    plotCI(stats$mean, dummy, li=stats$lower, ui=stats$upper, err="x", xlim=c(-11,46), axes=F, ann=T, ylab= "", xlab= "difference in predictive accuracy [%]", lwd=1.5 )
    axis(side = 2, at = c(1:length(dummy)), labels= names(dummy), lwd=0.1)
    axis(side = 1, lwd=0.1, lwd.ticks=1, at=c(-10,-5,0,5,10,15,20,25,30,35,40,45))
    box(lwd=1)
    abline(v=0, lty=2)
    abline(h=5.5, lty=1, lwd=0.5)
    stats <- cstats(rc, alg)
    plotCI(stats$mean, dummy, li= stats$lower, ui= stats$upper, err="x", xlim=c(-110,210), axes=F, ann=T, ylab= "", xlab= "difference in complexity [%]", lwd=1.5 )
    axis(side = 1, lwd=0.1, lwd.ticks=1, at=c(-150,-100,-50,0,50,100,150,200))
    box(lwd=1)
    abline(v=0, lty=2)
    abline(h=5.5, lty=1, lwd=0.5)
}
@

@@

     if( noise > 0){
        flipclasslist <- sample(totalpoints, totalpoints * (noise / 100), replace = F)

>>>>

\begin{table}
\begin{center}
\begin{tabular}{lrrrrrr}
\toprule
\textsc{Data-set} & \textsc{Instances} &  \multicolumn{5}{l}{\textsc{Attributes}}   \\ 
                  &                    &  binary              & nominal          &  ordered      &  metric     & classes \\ 
\midrule
Glass Identification$^\#$  &   214   &  - &  - &  - & 9 &  6      	 \\
Statlog Heart*         &   270   &  3 &   3 & 1 & 6 &   2     	 \\
Ionosphere$^\#$ &   351   &  2 &  - &  - & 32 & 2  	             \\
Musk$^+$  &  476  & - &  - &  - & 166 & 2                     	 \\
Breast Cancer Database$^\#$ &   685   &  - &  4 & 5 &  - &  2  	 \\
Pima Indians Diabetes$^\#$ &   768   &  - &   - & - & 8 &  2      	 \\
Vowel$^\#$  &  990  & - &  1 &  - & 9 & 11     	                     \\
Statlog German Credit* &  1000   &  2 &  10 & 1 & 7 &  2      	 \\
Contraceptive Method* &   1437   & 3 &  - &  4 &  2 &  3  	     \\
DNA$^\#$  &  3186  & 180 &  - &  - & - & 3   	                       \\
Spam$^+$  &  4601  & - &  - &  - & 57 & 2                      	 \\
MAGIC Gamma Telescope*  &  19020  & - &  - &  - & 10 & 2     	   \\
Servo$^\#$ & 167 &  - & 4 & - &  -  &  -											   \\
Boston Housing$^\#$ & 506 & 1 & - & - &  12  & -                 \\
MEL0101$^\diamond$ &   875 &  1 & 4 & 1 &  108  & -              \\
HDG0202$^\diamond$ &  3933 &  1 & 7 & 1 &   46  & -              \\
HDG0502$^\diamond$ & 8251 &  1 & 7 & 1 &   91  & -               \\
\bottomrule
\end{tabular}
\end{center}
\caption{Description of the evaluated benchmark data-sets. The data-sets marked with a $(^*)$ originate from the UCI machine learning repository \citep{Frank+Asuncion:2010} and are made available in the \pkg{evtree} package. Data-sets marked with a $(^\#)$ and $(^{+})$ are from the \proglang{R} packages \pkg{mlbench} \citep{mlbench} and \pkg{kernlab} \citep{kernlab}. The $3$ real-world data-sets from the Austrian DRG system are marked with a $(^\diamond)$. } 
\label{tab:dataDescription}
\end{table}

@@

preprocess <- function(d, dname = "datasetname", isclassification = TRUE){
    if(isclassification)
        d[,1:3] <- 1-d[,1:3]
    d <- as.data.frame(d)
    colnames(d) <- c("evtree", "rpart", "ctree","evtree", "rpart", "ctree")
    d[,2] <- d[,2] / d[,1] *100
    d[,3] <- d[,3] / d[,1] *100
    d[,1] <- d[,1] / d[,1] *100
    d[,5] <- d[,5] / d[,4] *100
    d[,6] <- d[,6] / d[,4] *100
    d[,4] <- d[,4] / d[,4] *100
    x <- d[ ,1:3]
    y <- d[ ,4:6]
    x <- reshape(x, idvar="samp", times=names(x), timevar = "alg",varying= list(names(x)), direction="long")
    y <- reshape(y, idvar="samp", times=names(y), timevar = "alg",varying= list(names(y)), direction="long")
    x$ds <- dname
    y$ds <- dname
    return( cbind(x,y) )
}

r <- preprocess(d = rglass, dname = "Glass Identification", isclassification = TRUE)
r <- rbind(r, preprocess(d = rheart, dname = "Statlog Heart", isclassification = TRUE))
r <- rbind(r, preprocess(d = rionosphere, dname = "Ionosphere", isclassification = TRUE))
r <- rbind(r, preprocess(d = rmusk, dname = "Musk", isclassification = TRUE))
r <- rbind(r, preprocess(d = rbreastcancer, dname = "Breast Cancer Database", isclassification = TRUE))
r <- rbind(r, preprocess(d = rpima, dname = "Pima Indians Diabetes", isclassification = TRUE))
r <- rbind(r, preprocess(d = rvowel, dname = "Vowel", isclassification = TRUE))
r <- rbind(r, preprocess(d = rcredit, dname = "Statlog German Credit", isclassification = TRUE))
r <- rbind(r, preprocess(d = rcontraceptive, dname = "Contraceptive Method", isclassification = TRUE))
r <- rbind(r, preprocess(d = rdna, dname = "DNA", isclassification = TRUE))
r <- rbind(r, preprocess(d = rspam, dname = "Spam", isclassification = TRUE))
r <- rbind(r, preprocess(d = rmagicgamma, dname = "Magic Gamma Telescope", isclassification = TRUE))
r <- rbind(r, preprocess(d = rservo, dname = "Servo", isclassification = FALSE))
r <- rbind(r, preprocess(d = rbostonhousing, dname = "Boston Housing", isclassification = FALSE))
r <- rbind(r, preprocess(d = rmel0101, dname = "MEL0101", isclassification = FALSE))
r <- rbind(r, preprocess(d = rhdg0202, dname = "HDG0202", isclassification = FALSE))
r <- rbind(r, preprocess(d = rhdg0502, dname = "HDG0502", isclassification = FALSE))

@@

plotd <- function(alg){
    par(las=1, cex=1.5, cex.lab=1.3, mfrow= c(1,2), omd=c(0.16,1,0,1), mar=c(5, 0, 1, 2) + 0.1)
    dummy <- c(1:17)
    names(dummy) <- uniquedatasetlabels
    stats <- cstats(ra, alg)
    plotCI(stats$mean, dummy, li=stats$lower, ui=stats$upper, err="x", xlim=c(-11,46), axes=FALSE, ann=TRUE, ylab= "", xlab= "difference in predictive accuracy [%]", lwd=1.5 )
    axis(side = 2, at = c(1:length(dummy)), labels= names(dummy), lwd=0.1)
    axis(side = 1, lwd=0.1, lwd.ticks=1, at=c(-10,-5,0,5,10,15,20,25,30,35,40,45))
    box(lwd=1)
    abline(v=0, lty=2)
    abline(h=5.5, lty=1, lwd=0.5)
    stats <- cstats(rc, alg)
    plotCI(stats$mean, dummy, li= stats$lower, ui= stats$upper, err="x", xlim=c(-110,210), axes=FALSE, ann=TRUE, ylab= "", xlab= "difference in complexity [%]", lwd=1.5 )
    axis(side = 1, lwd=0.1, lwd.ticks=1, at=c(-150,-100,-50,0,50,100,150,200))
    box(lwd=1)
    abline(v=0, lty=2)
    abline(h=5.5, lty=1, lwd=0.5)
}
@

@@

     if( noise > 0){
        flipclasslist <- sample(totalpoints, totalpoints * (noise / 100), replace = FALSE)
